Bayesian RKHS-based methods in functional regression\support{Support information of the article.}

### 2022-03-15 paper/paper.tex
Add some comments, pending revision

\begin{comment}
  Quizás falta comentar algo de los resultados experimentales...
\end{comment}

@@

There are at least three alternative, equivalent ways of defining the RKHS associated with \(K\), and each of them gives a different insight into the relationship between the process itself and the underlying RKHS.

@@

When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get a sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\). Assuming now a previously unseen test set \(\mathcal D_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples:

@@

Another thing to consider is that in this case we are approximating a probability, so before producing an output we need to transform the predicted values to \(\{0, 1\}\). We will follow the standard convention of considering a threshold \(\gamma = 0.5\) and predicting \(\hat Y=1\) if and only if \(\mathbb P(Y=1\mid X) \geq \gamma\). With this thresholding strategy in mind, the point estimation approach is analogous to the linear regression case:
\[
\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
\]
where \(\mathbb I\) is the indicator function and the hat estimates are obtained through a summary statistic of the corresponding marginal posterior distributions. On the other hand, the distributional estimates are somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples:
\begin{description}
  \item[Average sampled probability] If we choose to average the sampled probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 \mid X_i,\theta^{(m)*})\), we get the following estimator:
  \[
    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
  \]
  \item[Sampled majority vote] If we choose to average the sampled binary responses \(Y^{(m)*}\) instead, we compute the predictions as
\[
  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n',
\]
where each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
\end{description}
Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a multivariate binary classification procedure.

@@

As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments can be made:
\begin{itemize}
  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation). Our numerical studies showed that the results were much better with this choice than, say, with a random or null vector.
  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator. That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\) and \(\tilde\eta > 0\) is the actual tuning parameter. This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
\end{itemize}

To conclude this section, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away. For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions. Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}). In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(Y^*)\leq T(Y)\mid Y)\), and are computed by simply measuring the proportion of generated samples \(\{T(Y^{(m)*})\}\) that fall below the real value of the statistic. They are expected to be around 0.5 when the model accurately represents the data \citep[see][Ch.~6]{gelman1995bayesian}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supplementary Material, if any, should   %%
%% be provided in {supplement} environment  %%
%% with title and short description.        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{supplement}
\stitle{Title of Supplement A}
\sdescription{Short description of Supplement A, with DOI and/or links to additional material (code, ...)}

>>>>

\begin{comment}
  Quizás falta comentar algo de los resultados experimentales... también habría que hacer más hincapié en cuál es nuestra propuesta; que se vea bien.
\end{comment}

@@

There are at least three alternative, equivalent ways of defining the RKHS associated with \(K\), and each of them gives a different insight into the relationship between the process itself and the underlying RKHS:

@@

When performing the empirical posterior approximation, on each of the \(M\) steps of the iterative MCMC algorithm we get a sample \(\theta^{(m)*}\) of the posterior distribution \(\pi(\theta\mid \mathcal D_n)\). Assuming now a previously unseen test set \(\mathcal D'_{n'}\) in the same conditions as \(\mathcal D_n\), we propose to construct three different kinds of estimators based on the MCMC samples:

@@

Another thing to consider is that in this case we are approximating a probability, so before producing an output we need to transform the predicted values to \(\{0, 1\}\). We will follow the standard convention of considering a threshold \(\gamma = 0.5\) and predicting \(\hat Y=1\) if and only if \(\mathbb P(Y=1\mid X) \geq \gamma\). With this thresholding strategy in mind, the point estimation approach is analogous to the linear regression case:
\[
\hat Y_i = \mathbb I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
\]
where \(\mathbb I\) is the indicator function. The hat estimates are again obtained through a summary statistic of the corresponding marginal posterior distributions.

On the other hand, the distributional estimates are somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples:
\begin{description}
  \item[Average sampled probability] If we choose to average the sampled probabilities \(p_i^{(m)*} = \mathbb P(Y_i =1 \mid X_i,\theta^{(m)*})\), we get the following estimator:
  \[
    \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M p_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n'.
  \]
  \item[Sampled majority vote] If we choose to average the sampled binary responses \(Y^{(m)*}\) instead, we compute the predictions as
\[
  \hat Y_i = \mathbb I\left(\frac{1}{M} \sum_{m=1}^M Y_i^{(m)*} \geq \gamma\right), \quad i=1,\dots,n',
\]
where each \(Y_i^{(m)*}\) follows a Bernoulli distribution of parameter \(p_i^{(m)*}\), for \(m=1,\dots,M\).
\end{description}

Lastly, the variable selection method is essentially the same as in the case of linear regression: we select \(p\) times from each trajectory based on the posterior distribution of \(\tau\mid \mathcal D_n\), and then feed the reduced dataset to a multivariate binary classification procedure.

@@

As for the default values of the rest of hyperparameters in~\eqref{eq:prior-linear}, several comments can be made:
\begin{itemize}
  \item For the expected value \(b_0\) we propose to use (an approximation of) the MLE of \(\beta\) (see Section~\ref{sec:results} for details on the implementation) \incomment{La función de verosimilitud es complicada... pero nos basta con una aproximación como punto de partida}. Our numerical studies showed that the results were much better with this choice than, say, with a random or null vector.
  \item We found that the parameter \(g\) does not have as much influence on the final result, and the experimentation suggests that \(g=5\) is a good value.
  \item Lastly, we observed that the choice of \(\eta\) can have a considerable impact on the final estimator. That is why, in an effort to normalize its scale, we consider a compound parameter \(\eta = \tilde \eta \lambda_{\max}(G_\tau)\), where \(\lambda_{\max}(G_\tau)\) is the largest eigenvalue of the matrix \(G_\tau\) and \(\tilde\eta > 0\) is the actual tuning parameter. This standardization technique has been used previously in the literature; see for example \citet{grollemund2019bayesian}.
\end{itemize}

To conclude this section, it is worth mentioning that the Bayesian aspects of our model allow us to perform some model validation checks straight away. For example, we can derive credible intervals for each of the parameters, and in the case of linear regression, we can use the sampled values of \(\sigma^2\) as a measure of the uncertainty of the predictions. Moreover, we can perform various \textit{visual checks}, such as a plot comparing both the observed and posterior predictive distribution of the responses, or a diagram showing the distribution of some statistic \(T(Y^*)\) of the posterior predictive distribution (see Figure~\ref{fig:ppc}) \incomment{conseguir mejor calidad de la figura}. In addition, we can calculate the so-called \textit{Bayesian p-values} for several statistics, which are defined as \(P(T(Y^*)\leq T(Y)\mid Y)\), and are computed by simply measuring the proportion of generated samples \(\{T(Y^{(m)*})\}\) that fall below the real value of the statistic. They are expected to be around 0.5 when the model accurately represents the data \citep[see][Ch.~6]{gelman1995bayesian}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supplementary Material, if any, should   %%
%% be provided in {supplement} environment  %%
%% with title and short description.        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{supplement}
\stitle{Title of Supplement A}
\sdescription{Short description of Supplement A, with DOI and/or links to additional material (code, ...)}
